Notes on Control Plane Inference

### 2021-03-31 inference.tex
polish

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
      & \mid & \havoc & \textit{Havoc} & (a,g,p,\iota) \\
      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
      & \mid & x & \textit{Action Data Variable} & (a^*) \\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table} & (p)\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
      & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
      & \mid & t.\apply() & \textit{Table Application} & (p)\\
      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
    & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
  dataplane programs indicated by $p \in \Prog$, and table-free commands
  indicated by $g \in \GCL$.}
\label{fig:grammar}
\end{figure}

@@

\paragraph{Commands}
Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$),
assumptions ($\assume b$), assertions ($\assert b$), table application
($t.\apply()$), lexically-scoped nondeterministic selection ($\choose \rho t
c$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).

@@

Conversely $\iota \in \Instr \subset \Cmd$ represents a program instrumented
with symbolic variables. A program $\iota$ cannot apply tables using
$t.\apply()$, but instead uses a havoc statement $\choose \rho t c$, to
introduce symbolic variables $\rho$ in $c$.

@@

\begin{figure}[H]
  \[\begin{array}{l>{\triangleq}cl}
  \multicolumn{3}{l}{\edenote{e} : \Pkt \to \Model \rightharpoonup \Value }\\
    \edenote{v}~\pkt~\mu &&
    v \\
    \edenote{h.f}~\pkt~\mu &&
    pkt(h.f)\\
    \edenote{\rho.\matches[i]}~\pkt~\mu &&
    \mu(\rho).\matches[i] \\
    \edenote{\rho.\action}~\pkt~\mu &&
    \mu(\rho).\action \\
    \edenote{\rho.\data[i]} &&
    \mu(\rho).\data[i]\\
    \edenote{e_1 \binop e_2}~\pkt~\mu&&
    \edenote{e_1}~\pkt~\mu \binop
    \edenote{e_2}~\pkt~\mu \\
    \edenote{\unop e}~\pkt~\mu &&
    \unop \left(\edenote{e}~\pkt~\mu\right)
  \end{array}
  \]
  \caption{Semantics of Expressions}
  \label{fig:sem-expr}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
    \denote{h.f := e}^\sigma~\mu~\pkt && \{(\mu, \pkt\{h \mapsto \edenote{e}~\pkt~\mu\})\} \\
    \denote{\assume b}^\sigma~\mu~\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma~\mu~\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \error & \textit{otherwise}
    \end{cases} \\
    \denote{\choose \rho t c}^\sigma~\mu~\pkt &&
    \{(\mu'\setminus \rho,\pkt')
    \mid (\mu',\pkt') \in \denote{c}^\sigma\,(\mu\{\rho\mapsto r\})\,\pkt, r \in \sigma(t)\}
    \\
    \denote{c_1;c_2}^\sigma~\mu~\pkt&&
    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt' \\
    \denote{c_1 \choice c_2}^\sigma~\mu~\pkt &&
    \denote{c_1}^\sigma~\mu~\pkt \cup \denote{c_2}^\sigma~\mu~\pkt \\
    \denote{t.\apply()}^\sigma~\mu~\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma~\mu~\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{(\mu,\pkt)\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
  \todo[inline]{$\Model$ then $\Pkt$ everywhere.}
\end{figure}

@@

\section{Eliminating Tables}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  (\choose \rho t c)[\sigma] && \textrm{\textcolor{red}{undefined}}\\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    \{\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    \}~\choice\\
    \multicolumn{1}{c}{\vdots}\\
    \choice~\{\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    \}\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Elimination}
  \label{fig:table-elim}
\end{figure}

@@

\section{Instrumenting Tables}

@@

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \rho.\matches := \havoc; \\
    \quad \rho.\action := \havoc;\\
    \quad \rho.\data := \havoc; \\
    \quad \assume |\rho.\matches| = |t.\keys|; \\
    \quad \assume \rho.\action < |t.\actions|; \\
    \quad \assume |\rho.\data| < |t.\action[\rho.\action].\params| \\
    \quad \assume \rho.\matches = t.\keys; \\
    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left\{\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right\}\\
    \} \choice \{ \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    \}
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation}
  \label{fig:table-instrument}
\end{figure}

@@

\begin{lemma}[Models Invariant over Actions]
  \label{lem:models-invariant-over-actions}
  For a closed action command $c$, $\mu,\mu'
  \in \Model$, $\sigma \in \Value \to
  \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
  \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
  \end{lemma}

\begin{proof}
  Proceed by induction over $c$.
  \begin{enumerate}[align=left]
  \item[$(c = h.f = e)$] Immediate.
  \item[$(c = c_1;c_2)$]
    By IH($c_1$) there exists some $\pkt''$ such that
    $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
    By IH($c_2$) there exists some $\pkt'$ such that
    $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
    The result follows by definition of $\denote{-}$.
    \end{enumerate}
\end{proof}

\begin{theorem}[Instrumentation Equivalence]
  \label{thm:instr-equiv}
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
  $\mu \in \Model$, and packet $\pkt \in \Pkt$, $\denote{c}^\sigma~\mu~\pkt =
  \denote{c[?]}^\sigma~\mu~\pkt$.
\end{theorem}

\begin{proof}
  Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general.
  \begin{enumerate}[align=left]
  \item[$(c = h.f := e)$]
    Let $\sigma$,$\mu$,$\pkt$ be given.
    By definition $h.f:=e[?] = h.f:=e$. The result follows.
  \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
  \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
  \item[$(c = \choose \rho t)$] \textcolor{red}{Admitted}.
  \item[$(c = c_1;c_2)$]
    Let $\sigma,\mu,\pkt$ be given.
    By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
    Compute as follows:
    \[\begin{array}{>{\displaystyle}ll}
    \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
    = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
    & \text{by definition of } [?]  \\
    = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
    & \text{by definition of }\denote{-} \\
    = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
    & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
    = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
    & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
    = \denote{c_1;c_2}^\sigma~\mu~\pkt
    & \text{by definition of } \denote{-}
    \end{array}
    \]
  \item[$(c = c_1 \choice c_2)$] Similar, by IHs.

  \item[$(c = t.\apply())$]
    Let $\sigma$,$\mu$,$\pkt$ be given.
    Let $r_1,\ldots,r_n = \sigma(t)$.
    There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
    \begin{enumerate}
    \item[\textsc{Hit}] In this case, there is some $r_i$ such that $\satisfy \sigma \pkt \mu r_i.\matches = t.\keys$. By the well-formedness assumption,
      we know that this $r_i$ is unique, and that we can write
      $t.\actions[r_i.\action]$ as $\lambda \vec x. c$, and $|r_i.\data| = |\vec x|$.

      Compute as follows:
      \[\begin{array}{>{\displaystyle}ll}
      \denote{t.\apply()}^\sigma~\mu~\pkt \\
      = \denote{c[r_i.data/\vec x]}^\sigma~\mu~\pkt & \text{by definition of} \models \text{, and the uniqueness of }r_i  \\
      \end{array}
      \]

      Now, we compute the denotation of $t.\apply()[?]$ to see that it is equivalent to the above.
      The definition of $t.\apply()[?]$ gives us two alternatives. Consider the second, first:
      \[\begin{array}{>{\displaystyle}ll}
        \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys;\SKIP}^\sigma~\mu~\pkt \\
        = \denote{\assume \forall \rho \in t. \rho.\matches \neq t.\keys}^\sigma~\mu~\pkt
        & \textrm{\textcolor{red}{Redo with defaults}} \\
        = \begin{cases}
          \{(\mu,\pkt)\}, & \satisfy\sigma\pkt\mu \rho.\matches \neq t.\keys \\
          \{\}, &\textit{otherwise}
        \end{cases}
        & \textrm{By definition of }\denote{-}\\
        =  \{\}
        & \textrm{by \textsc{Hit} assumption}
      \end{array}
      \]

      Now let's consider the first branch, i.e.
      \[
      \denote{\choose \rho t \left\{
        \begin{array}{>{\displaystyle}l}
          \assume \rho.\matches = t.\keys;\\
          \bigchoice_{i=0}^{|t.\action| -1}
          \left\{\begin{array}{l}
          \assume \rho.\action = i;\\
          t.\actions[i](\rho.\data)
          \end{array}\right\}
        \end{array}\right\}
      }^\sigma~\mu~\pkt\]

      Compute the following inner result, by definitions and set theory
      \[
      \begin{array}{>{\displaystyle}ll}
      \bigcup_{r \in \sigma(t)}\denote{\begin{array}{>{\displaystyle}l}
          \assume \rho.\matches = t.\keys;\\
          \bigchoice_{i=0}^{|t.\action| -1}
          \left\{\begin{array}{l}
          \assume \rho.\action = i;\\
          t.\actions[i](\rho.\data)
          \end{array}\right\}
      \end{array}}^\sigma~\mu\{\rho \mapsto r\}~\pkt\\
      = \bigcup_{
        \substack{r \in \sigma(t) \\
          (\mu',\pkt') \in \denote{\assume \rho.\matches = t.\keys}^\sigma\,(\mu\{\rho \mapsto r\})\,\pkt}}
      \bigcup_{i=0}^{|t.\action| -1}\denote{
       \begin{array}{l}
          \assume \rho.\action = i;\\
          t.\actions[i](\rho.\data)
      \end{array}}^\sigma~\mu'~\pkt'

      \\\\
      (\textrm{by \textsc{Hit}, \textsc{NoShadow}, and the definition of }\denote{-}) \\
      \\
      = \bigcup_{i=0}^{|t.\action| -1}\denote{
       \begin{array}{l}
          \assume \rho.\action = i;\\
          t.\actions[i](\rho.\data)
      \end{array}}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
      = \bigcup_{j=0}^{|t.\action| -1}\denote{
       \begin{array}{l}
          \assume r_i.\action = j;\\
          t.\actions[j](r_i.\data)
      \end{array}}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
      \textrm{\textcolor{red}{TODO LEMMA}} \\
      = \denote{t.\actions[r_i.\action](r_i.\data)}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
      = \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt \\
      \end{array}\]

      By Lemma~\ref{lem:models-invariant-over-actions}, there's some $\pkt'$ such that
      \[\{(\mu\{\rho \mapsto r_i\}, \pkt')\} =
      \denote{c[r_i.\data/\vec x]}^\sigma~(\mu\{\rho \mapsto r_i\})~\pkt
      \]

      which means that the result of the first branch is
      \[\{(\mu, \pkt')\} \]

      Since $\rho$ doesn't occur in $c$, then there's some $\mu'$ such that also
      \[\{\mu', \pkt'\} = \denote{c[r_i.\data/\vec x]}^\sigma~\mu~\pkt \]

      and $\mu = \mu'$ by Lemma~\ref{lem:models-invariant-over-actions}.

      The result follows.

    \item[\textsc{Miss}] In this case, there is no $r_i$ such that
      $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
      of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
        this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
      is precisely the result we get when executing
      $\denote{t.apply()}^\sigma~\mu~\pkt$.

    \end{enumerate}

\section{Ghost State}

@@

\section{Tables as nondeterministic choice}

>>>>

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
      & \mid & \havoc & \textit{Havoc} & (a,g,p,\iota) \\
      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
      & \mid & x & \textit{Action Data Variable} & (a^*) \\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table} & (p)\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
      & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
      & \mid & t.\apply() & \textit{Table Application} & (p)\\
      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
      & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
      & \mid & \{c\} & \textit{Scope} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
  dataplane programs indicated by $p \in \Prog$, and table-free commands
  indicated by $g \in \GCL$.}
\label{fig:grammar}
\end{figure}

@@

\paragraph{Commands}
Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$),
assumptions ($\assume b$), assertions ($\assert b$), table application
($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).

@@

Conversely $\iota \in \Instr \subset \Cmd$ represents a program instrumented
with symbolic variables. A program $\iota$ cannot apply tables using
$t.\apply()$, but instead uses havoc statements, to
introduce symbolic variables $\rho$ in $c$.

@@

\begin{figure}[H]
  \[\begin{array}{l>{\triangleq}cl}
  \multicolumn{3}{l}{\edenote{e} : \Model \to \Pkt \rightharpoonup \Value }\\
    \edenote{v}~\mu~\pkt &&
    v \\
    \edenote{\havoc_w}~\mu~\pkt && \{[n]_w \mid n \in \mathbb N\}\\
    \edenote{h.f}~\mu~\pkt &&
    pkt(h.f)\\
    \edenote{\rho.\matches[i]}~\mu~\pkt &&
    \mu(\rho).\matches[i] \\
    \edenote{\rho.\action}~\mu~\pkt &&
    \mu(\rho).\action \\
    \edenote{\rho.\data[i]} &&
    \mu(\rho).\data[i]\\
    \edenote{e_1 \binop e_2}~\mu~\pkt&&
    \edenote{e_1}~\mu~\pkt \binop
    \edenote{e_2}~\mu~\pkt \\
    \edenote{\unop e}~\mu~\pkt &&
    \unop \left(\edenote{e}~\mu~\pkt\right)
  \end{array}
  \]
  \caption{Semantics of Expressions}
  \label{fig:sem-expr}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{l >{\triangleq}cl}
  \multicolumn{3}{l}{\denote{c}^\sigma : \Model \to \Pkt \to \mathcal{P}(\Model \times \Pkt)_\error }\\
    \denote{h.f := e}^\sigma\,\mu\,\pkt && \{(\mu, \pkt\{h \mapsto v\} \mid v \in \edenote{e}\,\mu\,\pkt\})\} \\
    \denote{\assume b}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \{\} & \textit{otherwise} \\
    \end{cases}\\
    \denote{\assert b}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \{(\mu,\pkt)\} & \satisfy \sigma \pkt \mu b \\
      \error & \textit{otherwise}
    \end{cases} \\
    \denote{c_1;c_2}^\sigma\,\mu\,\pkt&&
    \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma\,\mu'\,\pkt' \\
    \denote{c_1 \choice c_2}^\sigma\,\mu\,\pkt &&
    \denote{c_1}^\sigma\,\mu\,\pkt \cup \denote{c_2}^\sigma\,\mu\,\pkt \\
    \denote{\{c\}}^\sigma\,\mu\,\pkt &&
    \{(\mu,\pkt') \mid (\mu',\pkt') \in \denote{c}^\sigma\,\mu\,\pkt \}\\
    \denote{t.\apply()}^\sigma\,\mu\,\pkt &&
    \begin{cases}
      \denote{c[\vec d/\vec x]}^\sigma\,\mu\,\pkt,
      & \text{where}~\lambda \vec x.~c = r.\action, \vec d = r.\data \\
      & \text{where}~r = \argmin_i\{\satisfy \sigma \pkt \mu b_i\} \\
      & \text{where}~b_i = \sigma(t.\id)[i].\matches = t.\keys \\
      \{(\mu,\pkt)\}, & \mathit{otherwise}
    \end{cases}
  \end{array}\]
  \caption{Semantics of Commands}
  \label{fig:com-sem}
\end{figure}

@@

\subsection{Eliminating Tables}

@@

\begin{figure}
  \[\begin{array}{r>{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[\sigma] : \Prog \to \GCL} \\
  (h.f := e)[\sigma] && h.f := e   \\
  (\assume b)[\sigma] && \assume b \\
  (\assert b)[\sigma] && \assert b \\
  (c_1;c_2)[\sigma] && c_1[\sigma];c_2[\sigma] \\
  (c_1 \choice c_2)[\sigma] && c_1[\sigma] \choice c_2[\sigma] \\
  (\{c\})[\sigma] && \{c[\sigma]\} \\
  (t.\apply())[\sigma] &&
  \begin{array}[t]{l}
    (\\
    \quad \assume {r_1.\matches = t.\keys};\\
    \quad t.\action[r_1.\action](r_1.\data)\\
    )~\choice\\
    \multicolumn{1}{c}{\vdots}\\
    \choice~(\\
    \quad \assume {r_n.\matches = t.\keys};\\
    \quad t.\action[r_n.\action](r_n.\data)\\
    )\\
    \textrm{where}~[r_1;\ldots;r_n] = \sigma(t.\id)
  \end{array}
  \end{array}
  \]
  \caption{Table Elimination}
  \label{fig:table-elim}
\end{figure}

@@

\subsection{Instrumenting Tables}

@@

For simplicity, We're going to assume that all actions in a given table have the
same number of action data fields, and each field is the same number of bits.
This can be realized manually.

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (\{c\})[?] && \{c[?]\}\\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \rho.\matches := \overrightarrow\havoc; \\
    \quad \rho.\action := \havoc;\\
    \quad \rho.\data := \overrightarrow\havoc; \\
    \quad \assume \rho.\action < |t.\actions|; \\
    \quad \assume \rho.\matches = t.\keys; \\
    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left(\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right)\\
    \} \choice ( \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    )
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation. Typing information is omitted for
    brevity; each $\havoc$ must have a appropriate bit-width annotation. For
    simplicity, we assume each action in a table has the same number (and types) of
    parameters.}
  \label{fig:table-instrument}
\end{figure}

@@

%% \begin{lemma}[Models Invariant over Actions]
%%   \label{lem:models-invariant-over-actions}
%%   For a closed action command $c$, $\mu,\mu'
%%   \in \Model$, $\sigma \in \Value \to
%%   \Row^*$, $\pkt \in \Pkt$, there exists some $\pkt'$
%%   \[\denote{c}^\sigma\,\mu\,\pkt = \{\mu,\pkt'\}\]
%%   \end{lemma}

%% \begin{proof}
%%   Proceed by induction over $c$.
%%   \begin{enumerate}[align=left]
%%   \item[$(c = h.f = e)$] Immediate.
%%   \item[$(c = c_1;c_2)$]
%%     By IH($c_1$) there exists some $\pkt''$ such that
%%     $\denote{c_1}^\sigma\,\mu\,\pkt = \{(\mu,\pkt'')\}$.
%%     By IH($c_2$) there exists some $\pkt'$ such that
%%     $\denote{c_2}^\sigma,\mu\,pkt'' = \{(\mu, \pkt')\}$.
%%     The result follows by definition of $\denote{-}$.
%%     \end{enumerate}
%% \end{proof}

\begin{samepage}
\begin{theorem}[Instrumentation Equivalence]
  \label{thm:instr-equiv}
  For every $c \in \Cmd$, well-formed state $\sigma : \Value \to \Row^*$, model
  $\mu \in \Model$, and packet $\pkt \in \Pkt$, \[\denote{c}^\sigma~\mu~\pkt = \denote{c[?];\assume \state\sigma_c}^\sigma~\mu~\pkt.\]
\end{theorem}
\end{samepage}

\begin{proof}
  \todo[inline]{Proceed by induction on the structure of $c$, leaving $\sigma$, $\mu$, $\pkt$ general}

  %% \begin{enumerate}[align=left]
  %% \item[$(c = h.f := e)$]
  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
  %%   By definition $h.f:=e[?] = h.f:=e$. The result follows.
  %% \item[$(c = \assume b)$] Similar. $[?]$ is no-op.
  %% \item[$(c = \assert b)$] Similar. $[?]$ is no-op.
  %% \item[$(c = c_1;c_2)$]
  %%   Let $\sigma,\mu,\pkt$ be given.
  %%   By definition $(c_1;c_2)[?] = c_1[?]; c_2[?]$.
  %%   Compute as follows:
  %%   \[\begin{array}{>{\displaystyle}ll}
  %%   \denote{(c_1;c_2)[?]}^\sigma~\mu~\pkt \\
  %%   = \denote{c_1[?];c_2[?]}^\sigma~\mu\pkt
  %%   & \text{by definition of } [?]  \\
  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1[?]}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
  %%   & \text{by definition of }\denote{-} \\
  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2[?]}^\sigma~\mu'~\pkt'
  %%   & \text{by } \mathit{IH}(c_1, \mu, \pkt) \\
  %%   = \bigcup_{(\mu',\pkt') \in \denote{c_1}^\sigma\,\mu\,\pkt}\denote{c_2}^\sigma~\mu'~\pkt'
  %%   & \text{by } \mathit{IH}(c_2, \mu', \pkt') \\
  %%   = \denote{c_1;c_2}^\sigma~\mu~\pkt
  %%   & \text{by definition of } \denote{-}
  %%   \end{array}
  %%   \]
  %% \item[$(c = c_1 \choice c_2)$] Similar, by IHs.

  %% \item[$(c = t.\apply())$]
  %%   Let $\sigma$,$\mu$,$\pkt$ be given.
  %%   Let $r_1,\ldots,r_n = \sigma(t)$.
  %%   There are two cases, either the packet hits a row, or hits no row. Explore each case separately.
  %%   \begin{enumerate}
  %%   \item[\textsc{Hit}]
  %%     \textitP{}

  %%   \item[\textsc{Miss}] In this case, there is no $r_i$ such that
  %%     $\satisfy\sigma\pkt\mu r_i.\matches = t.\keys$, which means that the LHS
  %%     of the toplevel choice operator should evaluate to $\{\}$ \todo{compute
  %%       this}, and the RHS evaluates to $\{\mu,\pkt\}$\todo{compute this}, which
  %%     is precisely the result we get when executing
  %%     $\denote{t.apply()}^\sigma~\mu~\pkt$.

  %%   \end{enumerate}

  %% \end{enumerate}
\end{proof}

\subsection{Ghost State}

@@

\subsection{Tables as nondeterministic choice}
